\chapter{Research questions}
\label{ch:research-questions}

\section{Main research question}
In mature open-source projects with documented security incidents, are detectable AI-assisted contributions associated with different technical debt proxies, review dynamics, and security-traceability patterns than non-AI-signaled contributions?

\section{Sub-questions}
\begin{enumerate}
    \item \textbf{RQ1: Detection.} To what extent are detectable AI-assisted contributions identifiable in the selected projects, and what is their observable prevalence over time?
    \item \textbf{RQ2: Technical debt and review.} How do AI-signaled and non-AI-signaled contributions differ in technical debt proxies and review indicators, and how do those indicators differ between the pre-2022 and post-2022 periods?
    \item \textbf{RQ3: Security traceability.} Can documented vulnerabilities be traced to contribution and review pathways involving AI-signaled code, and what observable review characteristics accompany those pathways?
    \item \textbf{RQ4: Public discourse.} How do maintainers and contributors describe and frame the effects of AI-assisted contributions on review burden, quality control, and security in public repository discussion?
\end{enumerate}

\section{Analytical flow}
The analytical flow of the thesis is sequential. RQ1 establishes what can be observed by constructing a labelled dataset of detectable AI-assisted pull requests. RQ2 then compares AI-signaled and non-AI-signaled contributions using maintainability and review indicators, while also placing the post-2022 period against a pre-2022 baseline. RQ3 narrows the focus to security-relevant traceability by linking published vulnerabilities to fix commits, pull requests, and their observable review characteristics. RQ4 interprets these quantitative patterns through public repository discourse, showing how maintainers and contributors describe, justify, or problematize AI-assisted development in practice.
